\textbf{Appendix}\\ Bound within Boundaries: How Well Do Protected Areas Match Movement Corridors of Their Most Mobile Protected Species?

### 2021-01-18 Chapter_0/04_Manuscript/Manuscript.tex
Revision No. 2 implemented

\subsection{Least-Cost Paths and Corridors}
To identify potential movement corridors of dispersing wild dogs, we specified
source points and calculated factorial least-cost paths (LCPs) as well as
factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To specify
source points, we applied the omnidirectional \textit{go-through} approach
proposed by \citep{Koen.2014} and placed 68 regularly spaced source points along
the map border. This approach has been shown to reduce potential biases arising
from the selection of inappropriate source locations \citep{Koen.2014,
Pitman.2017}. To assess the sensitivity of our results with respect to the
location of source points, we also applied an alternative method and placed 68
source points within protected areas that are large enough to sustain viable
wild dog populations (further details and corresponding results in Appendix
A.9). In either case, the 68 source points resulted in 2'278 unique pairwise
combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
LCPs and LCCs between source points using the R-package \textit{gdistance}
(details in Appendix A.6). After computation, we tallied overlapping
LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
discrete paths, whereas LCCs return continuous corridors, we present both
methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. Significance codes: ** \(p
    < 0.05\), *** \(p < 0.01\).(b) Results from the k-fold cross validation for
    case-control studies. The left graph shows rank frequencies of
    \textit{realized} steps according to predictions, whereas the right graph
    shows rank frequencies of \textit{randomly selected} steps according to
    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
    resulting from 100 repetitions of the k-fold cross validation. The blue
    smoothing line was fitted using a locally weighted polynomial regression and
    serves to aid the eye in detecting the trends. Correlation coefficients
    suggest that our prediction was significant and robust, evidenced by the
    fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
    \(\bar{r}_{s, random}\) did overlap and by the fact that there was strong
    and significant correlation between ranks and associated frequency for
    realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

@@

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black semicircles along the map border) and
    corresponding least-cost paths between them. Continuous thin black lines
    indicate the borders of the KAZA-TFCA, whereas dashed black lines delineate
    country-borders. (b) Least-cost corridors between the same source points as
    illustrated in subfigure (a). For ease of spatial reference, we also labeled
    some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

@@

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersing wild dogs gain access to more peripheral regions of the
KAZA-TFCA. It appears that the absence of human activities, the central position
within the KAZA-TFCA, and the presence of relatively impermeable water bodies
(e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
highly frequented corridor. Furthermore, the lack of permeable areas between
peripheral source points often resulted in corridors stretching longer Euclidean
distances, in an arc-shaped route via a stretch of suitable habitat through the
Okavango-Linyanty ecosystem. This is an expected outcome in case structural and
functional connectivity coincide, i.e. when dispersers follow suitable habitats
to disperse \citep{Fattebert.2015, Hauenstein.2019}. The key role of the
Okavango-Linyanti region for overall connectivity within the KAZA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area, it has neither the status of a
National Park nor that of a Game Reserve. A similar case of non-formally
protected but key dispersal landscape is represented by the area south of Kafue
NP in Zambia, for which a disruption of its main and narrow dispersal corridor
would result in considerable isolation of its subpopulations. We also revealed a
potential southwards corridor between the Okavango-Linyanti ecosystem and the
Central Kalahari National Park. \cite{Elliot.2014} and \cite{Cushman.2018}
identified a similar corridor for dispersing lions, suggesting that upholding
and protecting a link between those ecosystems is pivotal. Some areas through
which the corridor runs are neither part of the KAZA-TFCA nor profit from any
form of protection status. Human presence and activities along the national road
that longitudinally traverses this corridor may limit functional connectivity
\citep{Cozzi.2020}.

>>>>

\subsection{Least-Cost Paths and Corridors}
To identify potential movement corridors of dispersing wild dogs, we specified
source points and calculated factorial least-cost paths (LCPs) as well as
factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To select
source points, we followed the omnidirectional \textit{go-through} approach
proposed by \cite{Koen.2014} and placed 68 regularly spaced source points along
the map border \citep{Koen.2014, Pitman.2017}. While this approach tends to
identify high connectivity towards the map boundaries, it reduces potential
biases caused by the selection of unreasonable source points \citep{Koen.2014}.
To assess the sensitivity of our results with respect to the location of source
points, we reran the same analysis using 68 source points located within
protected areas that are large enough to sustain viable wild dog populations
(further details and corresponding results in Appendix A.9). In either case, the
68 source points resulted in 2'278 unique pairwise combinations and therefore
2'278 unique LCPs and LCCs. We computed factorial LCPs and LCCs between source
points using the R-package \textit{gdistance} (details in Appendix A.6). After
computation, we tallied overlapping LCPs and LCCs, respectively, into single
connectivity maps. Because LCPs return discrete paths, whereas LCCs return
continuous corridors, we present both methods. R-codes showcasing the main steps
for sections 2.5 to 2.6 can be downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. \(ta\) = turning
    angle, \(sl\) = step length. Whiskers delineate the 95\%-CIs for estimated
    parameters. Significance codes: ** \(p < 0.05\), *** \(p < 0.01\). (b)
    Results from the k-fold cross validation for case-control studies. The left
    graph shows rank frequencies of \textit{realized} steps according to
    predictions, whereas the right graph shows rank frequencies of
    \textit{randomly selected} steps according to predictions. \(\bar{r}_s\)
    indicates the mean correlation coefficient resulting from 100 repetitions of
    the k-fold cross validation. The blue smoothing line was fitted using a
    locally weighted polynomial regression and serves to aid the eye in
    detecting the trends. Correlation coefficients suggest that our prediction
    was significant and robust, evidenced by the fact that the confidence
    intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did
    overlap and by the fact that there was strong and significant correlation
    between ranks and associated frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

@@

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black semicircles along the map border) and
    corresponding least-cost paths between them. Continuous black lines indicate
    the borders of the KAZA-TFCA, whereas dashed black lines delineate
    country-borders. (b) Least-cost corridors between the same source points as
    illustrated in subfigure (a). For ease of spatial reference, we also labeled
    some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

@@

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersing wild dogs gain access to more peripheral regions of the
KAZA-TFCA. It appears that the absence of human activities, the central position
within the KAZA-TFCA, and the presence of relatively impermeable water bodies
(e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
highly frequented corridor. Furthermore, the lack of permeable areas between
peripheral source points often resulted in corridors stretching longer Euclidean
distances, in an arc-shaped route via a stretch of suitable habitat through the
Okavango-Linyanty ecosystem. This is an expected outcome in case structural and
functional connectivity coincide, i.e. when dispersers follow suitable habitats
to disperse \citep{Fattebert.2015, Hauenstein.2019}. The key role of the
Okavango-Linyanti region for overall connectivity within the KAZA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area, it has neither the status of a
National Park nor that of a Game Reserve. A similar case of non-formally
protected but key dispersal landscape is represented by the area south of Kafue
NP in Zambia, for which a disruption of its main and narrow dispersal corridor
or high disperser mortality due for example to human persecution or vehicle
collision would result in considerable isolation of its subpopulations. We also
revealed a potential southwards corridor between the Okavango-Linyanti ecosystem
and the Central Kalahari National Park. \cite{Elliot.2014} and
\cite{Cushman.2018} identified a similar corridor for dispersing lions,
suggesting that upholding and protecting a link between those ecosystems is
pivotal. Some areas through which the corridor runs are neither part of the
KAZA-TFCA nor profit from any form of protection status. Human presence and
activities along the national road that longitudinally traverses this corridor
may limit functional connectivity \citep{Cozzi.2020}.

### 2021-01-18 Chapter_0/04_Manuscript/Manuscript_Support.tex
Revision No. 2 implemented

>>>>

%------------------------------------------------------------------------------
%  Appendix S8: Random Effects Variability
%------------------------------------------------------------------------------
\newpage
\section{Habitat Selection Model: Random Effects}
\begin{figure}[hbtp]
  \includegraphics[width = \textwidth]{99_RandomEffects}
  \caption{Plot of random effects showing variability across dispersal
  coalitions. The coalition ID refers to the individual in the dispersal
  coalition that was equipped with a GPS collar.}
  \label{RandomEffects}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S9: Alternative Source Points
%------------------------------------------------------------------------------
\newpage
\section{Alternative Source Points}
Because we were interested in assessing the sensitivity of our connectivity maps
with respect to the location of source points, we implemented two competing
approaches to select source points. In a first run, we applied an
omnidirectional \textit{go-through} approach as proposed by \cite{Koen.2014} and
applied by \cite{Pitman.2017}. That is, we distributed 68 equally spaced source
points along the map border and we calculated least-cost paths and least-cost
corridors for all pairwise combinations of them. In contrast to other methods,
this approach has been shown to reduce unwanted biases arising from the
selection of unreasonable source points. Furthermore, such an approach enables
the identification of dispersal routes that do not necessarily run between
occupied patches of suitable habitat. In a second run, we followed the approach
described in \cite{Elliot.2014} and overlayed the study area with a regular grid
of points spaced 100 km apart. From this regular grid we only considered those
points that fell within protected areas \(>\) 700 km\textsuperscript{2}, which
conforms with home-range requirements of African wild dogs \citep{Pomilia.2015}.
Finally, we defined centroids as source points for those protected areas \(>\)
700 km\textsuperscript{2} that were not assigned any source points from the
regular grid. Because wild dogs residing outside of protected areas are rare and
unviable \citep{VanDerMeer.2014}, we consider such a selection of source points
within protected areas to be reasonable. The second approach also resulted in 68
source points, implying 2'278 unique pairwise combinations and therefore 2'278
unique LCPs and LCCs per approach. \Cref{AlternativeSourcePoints} illustrates
the resulting least-cost paths and least-cost corridors and suggests that both
methods produce qualitatively similar results. In fact, map values of
\Cref{AlternativeSourcePoints} (b1) and (b2) correlate with \(r\) = 0.86.
Nevertheless, one can see that connectivity towards the map borders is more
pronounced using the first approach and that the final connectivity map in (b1)
appears to be a  ``stretched'' version of the map in (b2). This is an expected
result given that all source points were located at the map borders. Regardless
of the method, we find that the major dispersal corridors run within the borders
of the KAZA-TFCA and that mainly impenetrable landscape remains beyond its
borders.

\begin{figure}[hbtp]
  \begin{center}
  \includegraphics[width = \textwidth]{99_LeastCost_Suppl}
  \caption{Comparison of least-cost paths (a1 and a2) and least-cost corridors
  (b1 and b2) using different source points. In subfigures (a1) and (b1) 68
  source points were placed along the map border (black and white semicircles).
  In subfigures (a2) and (b2), on the other hand, source points (black and white
  circles) were placed in protected areas (gray) that were larger than 700
  km\textsuperscript{2}, which conforms to home range requirements for viable
  wild dog populations \citep{Pomilia.2015}. Continuous black lines indicate the
  borders of the KAZA-TFCA, whereas dashed black lines delineate
  country-borders. For ease of spatial reference, we also labeled some national
  parks (NPs, in dark-grey).}
  \label{AlternativeSourcePoints}
  \end{center}
\end{figure}
